Prediction of 2024 US election ...

### 2024-10-30 paper/paper.qmd
colin code update2

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
```

@@

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election.

@@

@@

cleaned_data |>
  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, answer) |>
  head(6) |>
  kable(
    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "answer"),
    booktabs = TRUE
    ) 
```

@@

```{r}
#| label: fig-clean-data-histogram-Harris-Trump
#| fig-cap: the average PCT vs State for Harris and Trump
#| echo: false
#| warning: false
#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
#| layout-nrow: 2
#| fig-width: 40
#| fig-height: 20
Harris_data = cleaned_data[cleaned_data$answer == "Harris", ]
Harris_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

Trump_data = cleaned_data[cleaned_data$answer == "Trump", ]
Trump_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

@@

>>>>

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(brms)
library(rstanarm)
```

@@

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.

In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election.

@@

## Explore the data

@@

cleaned_data |>
  select(pct, sample_size, pollscore, days_taken_from_election, state, methodology, candidate_name) |>
  head(6) |>
  kable(
    col.names = c("pct", "sample_size", "pollscore", "days_taken_from_election","state", "methodology", "candidate_name"),
    booktabs = TRUE
    ) 
```

@@

```{r}
#| label: fig-clean-data-histogram-Harris-Trump
#| fig-cap: the average PCT vs State for Harris and Trump
#| echo: false
#| warning: false
#| fig-subcap: ["average PCT vs State for Harris","PCT vs State for Trump"]
#| layout-nrow: 2
#| fig-width: 40
#| fig-height: 20
Harris_data = cleaned_data[cleaned_data$candidate_name == "Kamala Harris", ]
Harris_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

Trump_data = cleaned_data[cleaned_data$candidate_name == "Donald Trump", ]
Trump_data |>
  group_by(state) |>
  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
  ggplot(mapping = aes(x = state, y = avg_pct)) + 
  geom_col()  +
  theme_minimal() +
  labs(
    x = "State", y = "support percentage(pct)", title = "average support percentage(pct) for Harris per state"
  )

@@

## Basic Model
```{r}
just_harris_data = Harris_data |> na.omit()
model_MLR = lm(pct ~ pollscore + days_taken_from_election + methodology, data = just_harris_data)
summary(model_MLR)
just_harris_data <- just_harris_data |> mutate(fitted_value = predict(model_MLR),  num_harris = round((pct / 100) * sample_size, 0))

predictions = predict(model_MLR, just_harris_data)

ggplot(just_harris_data, aes(x = end_date)) +
  geom_point(aes(y = pct), color = "black") +
  geom_line(aes(y = fitted_value), color = "blue", linetype = "dotted") +
  facet_wrap(vars(methodology)) +
  theme_classic() +
  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
```

```{r}
baye_model_data = just_harris_data
baye_model_data$pct = as.factor(baye_model_data$pct)
baye_model_data$pollscore = as.factor(baye_model_data$pollscore)
baye_model_data$days_taken_from_election = as.factor(baye_model_data$days_taken_from_election)
baye_model_data$methodology = as.factor(baye_model_data$methodology)
baye_model_data$pct <- as.numeric(baye_model_data$pct)
# Define the Bayesian model with brms

formula <- pct ~ pollscore + days_taken_from_election + (1 | methodology) + (1 | state)

priors = normal(0, 2.5, autoscale = TRUE)

bayesian_model_1 <- stan_glmer(
  formula = formula,
  data = just_harris_data,
  family = gaussian(),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
pp_check(bayesian_model_1)
summary(bayesian_model_1)

# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)

```

```{r}

```

```{r}

# Transform Biden and Trump’s vote shares to fit a beta distribution
# Biden received 51.3%, Trump received 46.8%, totaling to approximately 98.1% (adjusted here for simplicity)
total_votes <- 100
biden_shape1 <- 51.3 / total_votes * 10
biden_shape2 <- (1 - 51.3 / total_votes) * 10
trump_shape1 <- 46.8 / total_votes * 10
trump_shape2 <- (1 - 46.8 / total_votes) * 10

# Set up priors using beta distributions based on 2020 vote shares
prior <- c(
  set_prior(paste0("beta(", biden_shape1, ", ", biden_shape2, ")"), class = "b", coef = "biden"),
  set_prior(paste0("beta(", trump_shape1, ", ", trump_shape2, ")"), class = "b", coef = "trump")
)

# Define the model formula
formula <- pct ~ pollscore + days_taken_from_election + methodology

# Run the Bayesian model
model <- brm(
  formula = formula,
  data = baye_model_data,
  family = categ0orical(),
  prior = prior,
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores =4
)

# Summarize the model
summary(model)

```
